---
title: "week1"
author: "Martin Haneferd"
date: "10 mars 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

Picked up at: https://www.r-bloggers.com/intro-to-text-analysis-with-r/
and
https://eight2late.wordpress.com/2015/05/27/a-gentle-introduction-to-text-mining-using-r/
and
http://handsondatascience.com/TextMiningO.pdf
and
https://cran.r-project.org/web/views/NaturalLanguageProcessing.html


## Load neccesary libraries

```{r, message=FALSE, cache=FALSE}
library(lsa) # cosine function
library(bigmemory)
library(slam)
library(Rstem)
library(sentiment)
library(plyr)
library(dplyr)
library(ggplot2)
library(wordcloud)
library(RColorBrewer)
library(SnowballC)
library(qdap)
library(qdapDictionaries)
library(scales)
library(Rgraphviz)
library(magrittr)
library(tm)
library(skmeans)
library(syuzhet) # Sentiment analyzis
# install.packages("tm.lexicon.GeneralInquirer", repos="http://datacube.wu.ac.at", type="source")
library(tm.lexicon.GeneralInquirer)
# install.packages("tm.plugin.sentiment", repos="http://R-Forge.R-project.org")
library(tm.plugin.sentiment) 

# Helpfunctions and variables:
remove_outliers <- function(x, na.rm = TRUE, ...) {
  qnt <- quantile(x, probs=c(0, .99), na.rm = na.rm, ...)
  H <- 1.5 * IQR(x, na.rm = na.rm)
  y <- x
  y[x < (qnt[1] - H)] <- NA
  y[x > (qnt[2] + H)] <- NA
  y
}

# Read stopwords:
myStopwords <- scan(file = "./stopwords_en.TXT", what = "character", sep ="\n", encoding = "UTF-8")

myStopwords <- c(myStopwords,"time","year","month","day","week","hour","minute") # Get rid of time variables
skipWords <- function(x){ removeWords(x, stopwords("english"))
      removeWords(x, myStopwords)
}

# Set seed so it can be reproduced
set.seed(123)
```

## Download and unpack data
```{r}
# Download the zipped files containing datasets
zipfile="./SwiftKey.zip"
if(!file.exists("./SwiftKey.zip")){
      fileUrl <- "https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip"
      download.file(fileUrl, destfile=zipfile, method="libcurl")
      # Unpack files to data directory
      unzip(zipfile)

}

# Look at the file sizes:
file.size(c("./final/en_US/en_US.blogs.TXT","./final/en_US/en_US.news.TXT","./final/en_US/en_US.twitter.TXT"))

# Just keep the english files. Remove the other languages, and the zipfile.
unlink(c("./final/de_DE","./final/fi_FI","./final/ru_RU"), recursive = TRUE, force = TRUE) 
unlink(zipfile)

# Clean up unneccesary values from environment:
rm(fileUrl)
rm(zipfile)
```

## Read the files
```{r, message=FALSE, cache=FALSE}

blogs <- scan(file = "./final/en_US/en_US.blogs.TXT", what = "character", sep ="\n", encoding = "UTF-8", skipNul=TRUE)

news <- scan(file = "./final/en_US/en_US.news.TXT", what = "character", sep ="\n", skipNul=TRUE)

twitter <- scan(file = "./final/en_US/en_US.twitter.TXT", what = "character", sep ="\n", skipNul=TRUE)

```

## Explore blogs data
There is some outlier far out, so I remove them in the histogram plot to get a better feeling of distribution of the major blog content.
```{r}
# Do some cleaning
blogs <- gsub("[^[:ascii:]]|\\’", "\'",blogs, perl=TRUE)
blogs <- tolower(blogs)
# Calculate the number of words.
numWords <- sapply(gregexpr("[[:alpha:]]+", blogs, perl=TRUE), function(x) sum(x > 0))
# Distribution of the numer of words in the Blogs:
summary(numWords) # See that there are some outliers!

numWords.cutted <- remove_outliers(numWords)
summary(numWords.cutted)
hist(numWords.cutted)

numChar <- sapply(blogs, nchar)
# Distribution of char's in the Blogs:
numChar.cutted <- remove_outliers(numChar) # Remove outliers
summary(numChar.cutted)
hist(numChar.cutted)
```
The majority of theblog posts have 10 or less words. I will remove the stop words and clean up the data, and group the blogs in 1-10 words, 11-20 words and 71-80 words, and see if there are any similarities of the top ten words.

### Cleaning the data
```{r}
b.docs <- Corpus(VectorSource(blogs))

tmFuncs <- list(stemDocument,removeNumbers,
                 removePunctuation, skipWords, stripWhitespace,
                 trimws)

# Clean and reduce to important words:
b.docs <- tm_map(b.docs, FUN = tm_reduce, tmFuns = tmFuncs)

# Transforming the tm corpus to vector of strings
blogs.reduce<-sapply(b.docs, as.character)
blogs.reduce[67] #Check
```

### Checking the distribution after reduction
The gap from 10'th to 20'th seems to be reduced. It looks more linear.
```{r}
# count words
numWords <- sapply(gregexpr("[[:alpha:]]+", blogs.reduce, perl=TRUE), function(x) sum(x > 0))

# remove the ones that only have 2 or less words
length(blogs.reduce) # Before
blogs.reduce <- blogs.reduce[numWords > 2]
length(blogs.reduce) #After 

# Do a re-count
numWords <- sapply(gregexpr("[[:alpha:]]+", blogs.reduce, perl=TRUE), function(x) sum(x > 0))

# Distribution of the numer of words in the Blogs after reduction:
numWords.cutted <- remove_outliers(numWords)
summary(numWords.cutted)
hist(numWords.cutted)

```

### Find the high frequencies of the words
```{r}
# Create av Document term matrix
b.dtm <- DocumentTermMatrix(b.docs)

# Find and plot the most used terms in the three sections of blogs
b.freq <- sort(col_sums(b.dtm), decreasing=TRUE)   
wf <- data.frame(word=names(head(b.freq, 15)), freq=head(b.freq, 15))

# Plot
ggplot(wf, aes(x = reorder(word, -freq), y = freq)) + 
      geom_bar(stat = "identity") +
      xlab("Terms (Word)") + 
      ylab("Frequency")
```

### Assosiactions in the blogs for the word love
There did not seem to be any associations to the word love, so I will not get into further investigation of that.
```{r}
# specifying a correlation limit of 0.2 (Not much)
b.assoc <- findAssocs(b.dtm, c("love"), corlimit=0.2) 
b.assoc
```
### Make a word cloud of the 50 most used terms
```{r}
wordcloud(names(head(b.freq, 50)), head(b.freq, 50), colors=brewer.pal(6, "Dark2"))
```
It looks that the majority of the blog posts is about everyday life; work, love, book, people, friends, family, school, hope, etc..

### Sentiment alayzis of the blogs
Lets 1st find out the number and distribution of positive and negative blogs.
```{r}
b.tdm <- TermDocumentMatrix(b.docs)
# Score the blogs in positive or negative
pos.score <- tm_term_score(b.tdm, terms_in_General_Inquirer_categories("Positiv")) # this lists each document with number below

neg.score <- tm_term_score(b.tdm, terms_in_General_Inquirer_categories("Negativ")) 

df.scores <- data.frame(positive = pos.score, negative = neg.score)
df.scores <- transform(df.scores, net = positive - negative)
# Find the spread of the scores
summary(df.scores$net)

# Look at the histogram of the pos/neg blogs
hist(df.scores$net, breaks = 100)
# Looks like there are more negative blogs than positive according to the histogram plot, but check this further

# Find three most positive and three most negative blogs
ord <- df.scores[ order(-df.scores[,3]), ]
most.positive <- as.numeric(rownames(head(ord,3)))
most.negative <- as.numeric(rownames(tail(ord,3)))

# Investigate what they are about:
# The negative:
blogs[most.negative]

# The positive:
blogs[most.positive]

# Count the positive, negative and neutral blogs
sum(  df.scores$net < 0  ) # number of negative blog posts
sum(  df.scores$net > 0  ) # number of positive
sum(  df.scores$net == 0  ) # Number of neutrals

# It is clear that there are more positive than negative blog posts. The histogram above must be adjusted for the neutrals in some ways.
```

And 2nd find out the mood of the blogs:
```{r}
b.sentiment <- get_nrc_sentiment(blogs)
b.df.sentiment <-data.frame(t(b.sentiment))

b.td_new <- data.frame(rowSums(b.df.sentiment))
#The function rowSums computes column sums across rows for each level of a grouping variable.
 
#Transformation and  cleaning
names(b.td_new)[1] <- "count"
b.td_new <- cbind("sentiment" = rownames(b.td_new), b.td_new)
rownames(b.td_new) <- NULL
b.td_new<-b.td_new[1:8,]

#Visualisation
ggplot(b.td_new, aes(x = reorder(sentiment, -count), y = count)) + 
      geom_bar(stat = "identity", aes(fill=sentiment)) +
      xlab("Terms (Word)") + 
      ylab("Frequency")
```
The mood of the blogs looks like it's filled of trust, anticipation and joy.

### Try clustering the data
I will try clustering the data

## Explore Twitter data
```{r}
# Find number of words in each tweet
numWords <- sapply(gregexpr("[[:alpha:]]+", twitter, perl=TRUE), function(x) sum(x > 0))

# Distribution of the numer of words in the tweets:
summary(numWords)
hist(numWords)

# Find the number of characters
numChar <- sapply(twitter, nchar)
# Distribution of char's in the Tweet:
summary(numChar)
hist(numChar)


```

## Extra exploration of the twitter data for the Quiz week 1
```{r}
# Divide the number of lines where the word "love" (all lowercase) occurs by the number of lines the word "hate" (all lowercase) occurs
twitter.low <- tolower(twitter)
sum.love <- sum(grepl("\\<love\\>", twitter.low))
sum.hate <- sum(grepl("\\<hate\\>", twitter.low))
sum.love / sum.hate

# The one tweet in the en_US twitter data set that matches the word "biostats" says what?
twitter.low[grepl("biostats", twitter.low)]

#How many tweets have the exact characters "A computer once beat me at chess, but it was no match for me at kickboxing". (I.e. the line matches those characters exactly.)
sum(grepl("\\<A computer once beat me at chess, but it was no match for me at kickboxing\\>", twitter))
```

## Explore News data
```{r}
# Find number of words in each News
numWords <- sapply(gregexpr("[[:alpha:]]+", news, perl=TRUE), function(x) sum(x > 0))

# Distribution of the numer of words in the News:
summary(numWords)
plot(numWords)

# Find the number of characters
numChar <- sapply(twitter, nchar)
# Distribution of char's in the News:
summary(numChar)
plot(numChar)


```


## Cluster
```{r, cache=FALSE,message=FALSE}
####################
#####3
####### Cluster
#######
#################

# This sparse percentage denotes the proportion of empty elements. A sparse parameter of 0.7 means that we select from those terms which are less than 70% empty. We set the sparsity at 0.95 so that terms with at least that sparse percentage will be removed. So the terms we accept can be very empty- at most 95% empty. Then we can coerce the TD matrix into a regular matrix.
tdm <- TermDocumentMatrix(docs, control=list(wordLengths=c(3, 15), bounds = list(global = c(1000,10000))))

tdms <- removeSparseTerms(tdm, 0.99)
dim(tdms)
# Delete rows that are zero:
col_freqr <-  col_sums(tdms)
tdms <- tdms[,row_freqr>0]
dim(tdms)

set.seed(1000)
sam <- sample(seq(1, 12056), 12000)
tdms <- tdms[,sam]

memory.limit(size=60000)
m <- as.matrix(tdms)

memory.limit(size=60000)
#rownames(m) <- 1:nrow(m)

## hierarchical clustering
library(proxy)
#compute distance between document vectors
d <- dist(m, method="cosine")

hc <- hclust(d, method="average")

plot(hc, hang=-1)

cl <- cutree(hc, 15)

table(cl)

findFreqTerms(tdms[cl==1], 5)


#run hierarchical clustering using Ward’s method
groups <- hclust(d,method="ward.D")
#plot dendogram, use hang to ensure that labels fall below tree
plot(groups, hang=-1)

rect.hclust(groups, k=5)




# dtm <- DocumentTermMatrix(docs, control=list(wordLengths=c(4, 15)))

# Here we have told R to include only those words that occur in  500 to 2000 documents. We have also enforced  lower and upper limit to length of the words included (between 4 and 20 characters).


# Inspect dtm
dtm

freqr <-  col_sums(dtm)
length(freqr)
#create sort order (asc)
ordr <- order(freqr,decreasing=TRUE)
#inspect most frequently occurring terms
freqr[head(ordr)]
# The output:
#    one   will   like    can   time   just 
# 136263 116009 111796 109971 108426 100612 
# Dous not tell much, lets see what the least output is:
#inspect least frequently occurring terms
freqr[tail(ordr)]

# keep only  words that occur >1000 times in all docs
#dtm2 <- dtm[,which( freqr > 30000)]
#dtm2
#freqr <-  col_sums(dtm2)

# Delete rows that are zero:
row_freqr <-  row_sums(dtm)
dtm <- dtm[row_freqr>0,]

# I will investigate the words freedom and sea a little bit further. To see the associations
# We can also find associations with a word, specifying a correlation limit

assosiations <- findAssocs(dtm, c("film","game"), corlimit=0.2)


d <- data.frame(word = colnames(dtm),freq=freqr)
dd <- head(d[order(d$freq,decreasing=TRUE),], 20)

barplot(d[1:10,]$freq, las = 2, names.arg = d[1:10,]$word, col ="lightblue", main ="Most frequent words", ylab = "Word frequencies")


##########################################
#############
#############    Clustering
#############
##########################################
## do tfxidf
dtm_tfxidf <- weightTfIdf(dtm)
inspect(dtm_tfxidf[1:10, 5001:5010])

## do document clustering

### k-means (this uses euclidean distance)
m <- as.matrix(dtm_tfxidf)
rownames(m) <- 1:nrow(m)

### don't forget to normalize the vectors so Euclidean makes sense
norm_eucl <- function(m) m/apply(m, MARGIN=1, FUN=function(x) sum(x^2)^.5)
m_norm <- norm_eucl(m)


### Remove NA's
m_norm[is.na(m_norm)] <- 0

### cluster into 10 clusters
cl <- kmeans(m_norm, 10)
cl

table(cl$cluster)

### show clusters using the first 2 principal components
plot(prcomp(m_norm)$x, col=cl$cl)

findFreqTerms(dtm_tfxidf[cl$cluster==1], 50)
inspect(reuters[which(cl$cluster==1)])


### this is going to take 4-ever (O(n^2))
memory.limit(size=40000)
d <- dist(m, method="cosine")
hc <- hclust(d, method="average")
plot(hc, hang=-1)

cl <- cutree(hc, 5)
table(cl)

findFreqTerms(dtm[cl==1], 50)




###################################
# Quantitive analyzis of text
words <- dtms %>% as.matrix %>% colnames %>% (function(x) x[nchar(x) < 20])

length(words)

head(words, 15)

summary(nchar(words))

table(nchar(words))

dist_tab(nchar(words))






###################################

freq <- rollup(tdm, 2, na.rm=TRUE, FUN = sum)
class(freq)

# Explore the document term matrix
sumfreq <- rowSums(as.matrix(freq))
length(sumfreq)







# Clustering

# clus: a skmeans object
# dtm: a Document Term Matrix
# first: eg. 10 most frequent words per cluster
# unique: if FALSE all words of the DTM will be used
#         if TRUE only cluster specific words will be used 



# result: List with words and frequency of words 
#         If unique = TRUE, only cluster specific words will be considered.
#         Words which occur in more than one cluster will be ignored.



mfrq_words_per_cluster <- function(clus, dtm, first = 10, unique = TRUE){
  if(!any(class(clus) == "skmeans")) return("clus must be an skmeans object")

  dtm <- as.simple_triplet_matrix(dtm)
  indM <- table(names(clus$cluster), clus$cluster) == 1 # generate bool matrix

  hfun <- function(ind, dtm){ # help function, summing up words
    if(is.null(dtm[ind, ]))  dtm[ind, ] else  col_sums(dtm[ind, ])
  }
  frqM <- apply(indM, 2, hfun, dtm = dtm)

  if(unique){
    # eliminate word which occur in several clusters
    frqM <- frqM[rowSums(frqM > 0) == 1, ] 
  }
  # export to list, order and take first x elements 
  res <- lapply(1:ncol(frqM), function(i, mat, first)
                head(sort(mat[, i], decreasing = TRUE), first),
                mat = frqM, first = first)

  names(res) <- paste0("CLUSTER_", 1:ncol(frqM))
  return(res)
}

# dtm3 has no rows with zero.

LDA(dtm, 30)

rownames(dtm3) <- paste0("Doc_", 1:dim(dtm3)[1])
clus <- skmeans(dtm3, 3)
# weightBin(x)
# Download CLUTO at:
# http://glaros.dtc.umn.edu/gkhome/cluto/cluto/download

file.create("cluto.file.test")
con <- file("cluto.file.test", open = "rw") 
writeCM(dtm3, con)
readCM(con, clabel = NULL)

close(con)

cluto <- doc2mat(dtm3)

clus <- skmeans(dtm3, method = "CLUTO", k = 4, control = list(nruns = 20))

mfrq_words_per_cluster(clus, dtm3)
mfrq_words_per_cluster(clus, dtm3, unique = FALSE)


```

```{r}

```


